---
title: "Charitable Error Results"
output: pdf_document
latex_engine: xelatex
---

```{r, echo=FALSE}
library('methods')
library('knitr')
library('ggplot2')

overview <- function(var) {
    cat(sprintf("* %s = %d\n", "N", length(var)))
    cat(sprintf("* %s = %0.2f\n", "Mean", mean(var)))
    cat(sprintf("* %s = %0.2f\n", "Median", median(var)))
    cat(sprintf("* %s = %0.3f\n", "SD", sd(var)))
    cat(sprintf("* %s = %0.2f-%0.2f\n", "Range", min(var), max(var)))
}

cov <- read.delim('statistics.dat', header=TRUE, sep="\t")

```

# Converting to Charitable errors:

That is, assume that any phoneme missing is probably one of the errors.

```{r}
cov$Charitable <- cov$Unobserved - cov$DistinctErrors 
err <- cov[cov$Charitable > 0,]
ok <- cov[cov$Charitable <= 0,]
```

# Number of Absent Phonemes.

- the number of phonemes for each language that are absent from the language's transcription (more absences means poorer coverage).

- Note that a phoneme with allophones is NOT considered missing if at least one of its allophones is present e.g. if we have
  "a(a:, a)" and "a:" is found but not "a", then this is still considered not missing.

```{r, echo=FALSE}
p <- ggplot(cov, aes(Charitable))
p <- p + geom_histogram(binwidth=1)
p <- p + ggtitle('Number of Absent Phonemes (CHARITABLE)')
p <- p + theme_classic()
print(p)

pdf('histogram-CHARITABLE-absent_phonemes.pdf')
print(p)
x <- dev.off()
```




## Summary Statistics

```{r}
overview(cov$Charitable)
```

## Languages that are most poorly described by the transcript:

```{r, echo=FALSE}
kable(
    head(cov[order(-cov$Charitable),])[c("Language", "Charitable")]
)
```

## Languages that are completely described by the transcript:

```{r, echo=FALSE}
kable(
    cov[cov$Charitable <= 0,][c("Language", "Charitable", "DistinctErrors")]
)
cat(sprintf(
    "%d of %d languages",   
    nrow(cov[cov$Charitable==0,]),
    nrow(cov)
))
```

\clearpage



