---
title: "Correlation Results"
output: pdf_document
latex_engine: xelatex 
---


```{r, echo=FALSE}
library('methods')
library('knitr')
library('ggplot2')

cov <- read.delim('statistics.dat', header=TRUE, sep="\t")
cov$PropAbsent <- (cov$Unobserved / cov$InventoryLength) * 100

get_corlabel <- function(x, y) {
    r <- cor.test(x, y, method="spearman")
    if (r$p.value <= 0.05) {
        sig <- '*'
    } else {
        sig <- ''
    }
    sprintf(
        "r=%0.2f, p=%0.2f%s",
        r$estimate,
        r$p.value,
        sig
    )
}

```


# Do languages with a bigger inventory have more absent?

Languages with more phonemes have a much much higher rate of missing phonemes. 

```{r, echo=FALSE}
p <- ggplot(cov, aes(x=Unobserved, y=InventoryLength))
p <- p + geom_point()
p <- p + geom_smooth(method="glm")
p <- p + ggtitle(
    sprintf("Absent Phonemes vs Inventory Size (%s)",
        get_corlabel(cov$InventoryLength, cov$Unobserved)
    )
)
p <- p + xlab('Absent Phonemes') + ylab("Inventory Size")
p <- p + theme_classic()

print(p)

pdf('scatter-absences_vs_inventory_size.pdf')
print(p)
x <- dev.off()

cor.test(cov$Unobserved, cov$InventoryLength, method="spearman")
```

## Faceted by family

```{r, echo=FALSE}
t <- table(cov$Family)
families <- names(t[t> 1])
covs <- cov[cov$Family %in% families,]

p <- ggplot(covs, aes(x=Unobserved, y=InventoryLength))
p <- p + geom_point()
p <- p + geom_smooth(method="glm")
p <- p + facet_wrap(~Family)
p <- p + xlab('Absent Phonemes') + ylab("Inventory Size")
p <- p + theme_classic()

print(p)

pdf('scatter-absences_vs_inventory_size-faceted.pdf')
print(p)
x <- dev.off()
```


\clearpage


# Does coverage get better with a longer transcript?

- Coverage is not really getting much better as transcripts get longer. 
- Only a very weak and non-significant effect.
- Note the log scale on y axis as transcript length spans an order of magnitude.

```{r, echo=FALSE}
p <- ggplot(cov, aes(x=Unobserved, y=TranscriptLength))
p <- p + geom_point()
p <- p + scale_y_log10()
p <- p + geom_smooth(method="glm")
p <- p + ggtitle(
    sprintf("Absent Phonemes vs Transcript Length (%s)",
        get_corlabel(cov$TranscriptLength, cov$Unobserved)
    )
)
p <- p + xlab('Absent Phonemes') + ylab("Transcript Length")
p <- p + theme_classic()

print(p)

pdf('scatter-absences_vs_transcript_length.pdf')
print(p)
x <- dev.off()


cor.test(cov$Unobserved, cov$TranscriptLength, method="spearman")

```

\clearpage

```{r}
library(RColorBrewer)
colorCount = length(unique(cov$Family))
getPalette = colorRampPalette(brewer.pal(9, "Set1"))
 
p <- ggplot(cov, aes(x=Unobserved, y=InventoryLength, color=Family))
p <- p + geom_jitter()
p <- p + scale_color_manual(values = getPalette(colorCount))
p <- p + ggtitle("Absent Phonemes vs Inventory Size")
p <- p + xlab('Absent Phonemes') + ylab("Inventory Size")
p <- p + theme_classic()
p <- p + theme(legend.position="bottom")
#p <- p + guides(fill=guide_legend(nrow=2))

plot(p)

pdf('scatter-absences_vs_inventory_length-colored.pdf')
print(p)
x <- dev.off()
```

```{r}
p <- ggplot(cov, aes(x=Unobserved, y=TranscriptLength, color=Family))
p <- p + geom_jitter()
p <- p + scale_color_manual(values = getPalette(colorCount))
p <- p + ggtitle("Absent Phonemes vs Transcript Length")
p <- p + xlab('Absent Phonemes') + ylab("Transcript Length")
p <- p + theme_classic()
p <- p + theme(legend.position="bottom")
#p <- p + guides(fill=guide_legend(nrow=2))

plot(p) 

pdf('scatter-absences_vs_transcript_length-colored.pdf')
print(p)
x <- dev.off()
```
