---
title: "Coverage Results"
output: pdf_document
latex_engine: xelatex 
---


```{r, echo=FALSE}
library('methods')
library('knitr')
library('ggplot2')

cov <- read.delim('statistics.dat', header=TRUE, sep="\t")
cov$PropAbsent <- (cov$Unobserved / cov$InventoryLength) * 100

overview <- function(var) {
    cat(sprintf("* %s = %d\n", "N", length(var)))
    cat(sprintf("* %s = %0.2f\n", "Mean", mean(var)))
    cat(sprintf("* %s = %0.2f\n", "Median", median(var)))
    cat(sprintf("* %s = %0.3f\n", "SD", sd(var)))
    cat(sprintf("* %s = %0.2f-%0.2f\n", "Range", min(var), max(var)))
}

get_corlabel <- function(x, y) {
    r <- cor.test(x, y, method="spearman")
    if (r$p.value <= 0.05) {
        sig <- '*'
    } else {
        sig <- ''
    }
    sprintf(
        "r=%0.2f, p=%0.2f%s",
        r$estimate,
        r$p.value,
        sig
    )
}

```

# How big are the phoneme inventories?

```{r, echo=FALSE}
p <- ggplot(cov, aes(InventoryLength))
p <- p + geom_histogram(binwidth=1)
p <- p + ggtitle('Number of Phonemes in Inventory')
p <- p + xlab("Number of Phonemes")
p <- p + ylab("Number of Languages")
p <- p + theme_classic()

print(p)

pdf('histogram-phoneme_inventory_size.pdf')
print(p)
x <- dev.off()
```

\clearpage

## Summary Statistics

```{r}
overview(cov$InventoryLength)
```



## Biggest Phoneme Inventory

```{r, echo=FALSE}
cov.sorted <- cov[order(-cov$InventoryLength),]
cov.sorted <- cov.sorted[c("Language","InventoryLength")]
kable(head(cov.sorted))
```

## Smallest Phoneme Inventory

```{r, echo=FALSE}
kable(tail(cov.sorted))
```



\clearpage

# How big are the transcripts?

```{r, echo=FALSE}
p <- ggplot(cov, aes(TranscriptLength))
p <- p + geom_histogram(binwidth=20)
p <- p + ggtitle('Length of Transcript (Phonemes)')
p <- p + xlab("Number of Phonemes")
p <- p + ylab("Number of Languages")
p <- p + theme_classic()

print(p)

pdf('histogram-transcript_size.pdf')
print(p)
x <- dev.off()
```

\clearpage


## Summary Statistics

```{r}
overview(cov$TranscriptLength)
```


## Longest Transcript

```{r, echo=FALSE}
cov.sorted <- cov[order(-cov$TranscriptLength),]
cov.sorted <- cov.sorted[c("Language", "TranscriptLength")]
kable(head(cov.sorted))
```

## Shortest Transcript

```{r, echo=FALSE}
kable(tail(cov.sorted))
```


\clearpage


# Number of Absent Phonemes.

- the number of phonemes for each language that are absent from the language's transcription (more absences means poorer coverage).

- Note that a phoneme with allophones is NOT considered missing if at least one of its allophones is present e.g. if we have
  "a(a:, a)" and "a:" is found but not "a", then this is still considered not missing.

```{r, echo=FALSE}
p <- ggplot(cov, aes(Unobserved))
p <- p + geom_histogram(binwidth=1)
p <- p + ggtitle('Number of Unobserved Phonemes')
p <- p + theme_classic()
print(p)

pdf('histogram-absent_phonemes.pdf')
print(p)
x <- dev.off()
```

\clearpage

## Summary Statistics

```{r}
overview(cov$Unobserved)
```

## Languages that are most poorly described by the transcript:

```{r, echo=FALSE}
kable(
    head(cov[order(-cov$Unobserved),])[c("Language", "Unobserved")]
)
```

## Languages that are completely described by the transcript:

```{r, echo=FALSE}
kable(cov[cov$Unobserved==0,][c("Language", "Unobserved")])
cat(sprintf(
    "%d of %d languages",
    nrow(cov[cov$Unobserved==0,]),
    nrow(cov)
))
```

\clearpage



