---
title: "Results Notebook"
output: pdf_document
---

# Zipf

Read data

```{r}
z <- read.delim('results.dat', header=TRUE)
head(z)
```


## plot alpha parameter

```{r}
library(ggplot2)

p <- ggplot(z, aes(x=Alpha)) + geom_histogram() + xlim(0, 6) + theme_classic()
p <- p + geom_vline(xintercept=1, color="tomato")
p
```

Hmm. Not much around expected Zipf alpha of 1.

Summary of Alpha: 

```{r}
summary(z$Alpha)
sd(z$Alpha)
```



## compare models: Exponential

Negative means preference for the named model.

```{r}
p <- ggplot(z, aes(x=vsExp.R)) + geom_histogram() + xlim(-10, 10) + theme_classic()
p <- p + geom_vline(xintercept=0, color="gray")
p <- p + ggtitle('Exponential')
p
```

## compare models: Lognormal

Negative means preference for the named model.


```{r}
p <- ggplot(z, aes(x=vsLN.R)) + geom_histogram() + xlim(-10, 10) + theme_classic()
p <- p + geom_vline(xintercept=0, color="gray")
p
```

## compare models: Truncated Power law

```{r}
p <- ggplot(z, aes(x=vsTPL.R)) + geom_histogram() + xlim(-10, 10) + theme_classic()
p <- p + geom_vline(xintercept=0, color="gray")
p <- p + ggtitle('LogNormal')
p
```


## Combined plot:

```{r}
df <- data.frame(Language=z$Language, Model="Exponential", R=z$vsExp.R, P=z$vsExp.P)
df <- rbind(df, data.frame(
    Language=z$Language, Model="Log Normal", R=z$vsLN.R, P=z$vsLN.P    
))
df <- rbind(df, data.frame(
    Language=z$Language, Model="Truncated Power Law", R=z$vsTPL.R, P=z$vsTPL.P    
))
```

## P-values

```{r}
p <- ggplot(df, aes(x=P, group=Model, fill=Model))
p <- p + geom_histogram()
p <- p + facet_grid(Model~.)
p <- p + theme_classic()
p <- p + geom_vline(xintercept=0, color="gray")
p <- p + theme(legend.position="none")
p <- p + theme(
      strip.background = element_rect(colour=NA, fill=NA)
)
p
```

## Likelihood ratio

```{r}

p <- ggplot(df, aes(x=R, group=Model, fill=Model))
p <- p + geom_histogram() + xlim(-15, 5)
p <- p + facet_grid(Model~.)
p <- p + theme_classic()
p <- p + geom_vline(xintercept=0, color="gray")
p <- p + theme(legend.position="none")
p <- p + theme(
      strip.background = element_rect(colour=NA, fill=NA)
)
p
```


# Best fitting model

```{r}
BFM <- data.frame(Language=c(), Model=c(), R=c())
for (language in levels(df$Language)) {
    b <- data.frame(Language=c(language), Model=c('Power Law'), R=c(0))
    
    for (model in levels(df$Model)) {
        print(paste(language, model))
        m <- df[c(df$Language == language & df$Model == model),]
        if (m$P <= 0.05 & m$R < b$R) {
            b <- data.frame(Language=c(language), Model=c(model), R=c(m$R))
        }
    }
    BFM <- rbind(BFM, b)
}

p <- ggplot(BFM, aes(x=Model)) + geom_bar()

```

Best fitting models:
```{r}
print(summary(BFM$Model))
print(summary(BFM$Model) / length(BFM$Model))
```

Size of R:
```{r}
print(summary(BFM$R))
print(sd(BFM$R))

```




